---
output:
  html_document:
    df_print: kable
csl: ecology.csl
bibliography: refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = T, results = T, message = F, warning = F, eval = T, fig.dim = c(4,4))
library(dplyr)
library(drake)
library(multipanelfigure)
library(ggplot2)
theme_set(theme_bw())

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analyses", "caches", "all.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")


```


```{r load results and filter to desired temporal coverage}

desired_routes <- read.csv(here::here("analyses", "supporting_data", "ninetypercent_coverage_1988_2018.csv"), stringsAsFactors = F)

loadd(all_sims, cache=cache)

all_sims <- all_sims %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_preds_b, cache=cache)
all_preds_b <- all_preds_b %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_aics_b, cache=cache)
all_aics_b <- all_aics_b %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_preds_e, cache=cache)
all_preds_e <- all_preds_e %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_aics_e, cache=cache)
all_aics_e <- all_aics_e %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_cs_compares, cache =cache)
all_cs_compares <- all_cs_compares %>%
  filter(matssname %in% desired_routes$matssname)

```



```{r process biomass results}

# Select the best-fitting model for each route using AICc (simplest model with an AICc score within two units of the minimum)
# all_aics contains the AIC and AICc scores for every model fit to each route.

winning_aics <- all_aics_b %>%
  filter(model_family == "Gamma") %>% # I fit both Gamma and Gaussian models. The Gaussian are easier to interpret (slope terms can be read out directly) but are less strictly appropriate for strictly-positive biomass/energy use data.
  group_by(matssname, model_family) %>%
  mutate(minAICc = min(model_AICc)) %>% # identify best AICc
  mutate(deltaAICc = minAICc - model_AICc)  %>% # deltaAICc is the difference between each model's AICc and the best AICc for that pool
  filter(deltaAICc > -2) %>% # use 2 as a rule of thumb cutoff 
  arrange(matssname, modelcomplexity) %>% # rank by model complexity. 1 is simplest, 4 is most complex.
  mutate(rank = row_number()) %>% 
  filter(rank == 1) %>% # select the simplest model with deltaAICc within 2 of the best
  ungroup()


# Get predicted changes for winning models
# all_preds contains predicted trends for all models; the left_join effectively filters to the models in winning_aics.
winning_fits <- winning_aics %>%
  left_join(all_preds_b) 


# Get changes for winning models *only for routes for which there is some temporal trend*; i.e. the winning model is not the intercept-only model.
slope_winning_fits <- winning_fits %>%
  filter(modelcomplexity > 1)

# Pivot longer for plotting and summarizing
biomass_slopes_long <- slope_winning_fits %>%
  select(matssname, ratio_sim, ratio_real) %>%
  tidyr::pivot_longer(-matssname, names_to = "source", values_to = "biomass_ratio") %>%
  mutate(Dynamics = ifelse(source == "ratio_sim", "Individuals driven", "Observed"))

```


```{r process energy results}


winning_aics_energy <- all_aics_e %>%
  filter(model_family == "Gamma") %>% # I fit both Gamma and Gaussian models. The Gaussian are easier to interpret (slope terms can be read out directly) but are less strictly appropriate for strictly-positive biomass/energy use data.
  group_by(matssname, model_family) %>%
  mutate(minAICc = min(model_AICc)) %>% # identify best AICc
  mutate(deltaAICc = minAICc - model_AICc)  %>% # deltaAICc is the difference between each model's AICc and the best AICc for that pool
  filter(deltaAICc > -2) %>% # use 2 as a rule of thumb cutoff 
  arrange(matssname, modelcomplexity) %>% # rank by model complexity. 1 is simplest, 4 is most complex.
  mutate(rank = row_number()) %>% 
  filter(rank == 1) %>% # select the simplest model with deltaAICc within 2 of the best
  ungroup()


# Get predicted changes for winning models
# all_preds contains predicted trends for all models; the left_join effectively filters to the models in winning_aics.
winning_fits_energy <- winning_aics_energy %>%
  left_join(all_preds_e) 


# Get changes for winning models *only for routes for which there is some temporal trend*; i.e. the winning model is not the intercept-only model.
slope_winning_fits_energy <- winning_fits_energy %>%
  filter(modelcomplexity > 1)

# Pivot longer for plotting and summarizing
energy_slopes_long <- slope_winning_fits_energy %>%
  select(matssname, ratio_sim, ratio_real) %>%
  tidyr::pivot_longer(-matssname, names_to = "source", values_to = "energy_ratio") %>%
  mutate(Dynamics = ifelse(source == "ratio_sim", "Individuals driven", "Observed"))


# Combine energy and biomass results into one dataframe
winning_fits_both <- winning_fits %>%
  mutate(currency = "biomass") %>%
  bind_rows(mutate(winning_fits_energy, currency = "energy"))

```


```{r}

interaction_all_fits <- all_preds_b %>% 
  # get just the change ratios from the interaction models - regardless of whether the interaction model is the best-fitting model. this is because we want an estimate of the slope, we'll color by model fit in a minute
  group_by_all() %>% 
  filter(grepl( "* source",model_formula), model_family == "Gamma") %>%
  mutate(plus = substr(model_formula, 12,12) == "+") %>% filter(!plus) %>%
  ungroup() %>%
  # now join to *winning* formulas to get the best fitting formula
  left_join(select(winning_fits, matssname, model_formula, modelcomplexity) %>%
              rename(winning_formula = model_formula)) %>%
  select(winning_formula, ratio_sim, ratio_real, matssname, modelcomplexity) %>%
  mutate(currency = "Biomass")


interaction_all_fits_energy <- all_preds_e  %>% 
  # get just the change ratios from the interaction models - regardless of whether the interaction model is the best-fitting model. this is because we want an estimate of the slope, we'll color by model fit in a minute
  group_by_all() %>% 
  filter(grepl( "* source",model_formula), model_family == "Gamma") %>%
  mutate(plus = substr(model_formula, 12,12) == "+") %>% filter(!plus) %>%
  ungroup() %>%
  # now join to *winning* formulas to get the best fitting formula
  left_join(select(winning_fits_energy, matssname, model_formula, modelcomplexity) %>%
              rename(winning_formula = model_formula)) %>%
  select(winning_formula, ratio_sim, ratio_real, matssname, modelcomplexity) %>%
  mutate(currency = "Energy use")


all_interaction <- bind_rows(interaction_all_fits, interaction_all_fits_energy) %>%
  mutate(Syndrome = ifelse(modelcomplexity == 1, "No trend", ifelse(modelcomplexity == 2, "Coupled trend", "Decoupled trends")))

all_interaction_wide <- all_interaction %>%
  mutate(currency2 = gsub(" ", "_", currency)) %>%
  tidyr::pivot_wider( id_cols = matssname, names_from = currency2, values_from = c(ratio_sim, ratio_real, modelcomplexity, winning_formula, Syndrome)) %>%
  mutate(Same_syndrome = ifelse(Syndrome_Biomass == Syndrome_Energy_use,
                                "Same syndromes for both currencies",
                                "Different syndromes"))

```


## Example of biomass and energy showing different behaviors

For about 1/4 of routes, biomass and total energy use have different "syndromes" (coupled trend, decoupled trend, no trend):

```{r}

all_interaction_wide %>%
  group_by(Same_syndrome) %>%
  tally() %>%
  mutate(total = sum(n),
         prop = round(n/sum(n), digits = 2))

```


What stands out to me about this is that, in instances where they show different "syndromes" (left panel), usually the slope for energy use is less positive than for biomass (90% of the time for these routes; this is somewhat overrepresented compared to 78% of the time for routes with the same "syndrome"). 

```{r, fig.dim = c(6,4)}
ggplot(all_interaction_wide, aes(ratio_real_Biomass, ratio_real_Energy_use, color = Syndrome_Energy_use, shape = Syndrome_Biomass)) + 
  geom_point() + 
  facet_wrap(vars(Same_syndrome)) + 
  scale_x_log10() + 
  scale_y_log10() + 
  geom_abline(slope = 1, intercept = 0) + 
  geom_vline(xintercept = 1) + 
  geom_hline(yintercept = 1) +
  theme(legend.position = "bottom") +
  scale_color_viridis_d(end = .8) +
  xlab("Biomass\nRatio last:first fitted value") +
  ylab("Energy use\nRatio last:first fitted value")
# 
# all_interaction_wide %>%
#   filter(Same_syndrome == "Different syndromes") %>%
#   group_by(Syndrome_Biomass, Syndrome_Energy_use) %>%
#   tally() %>% 
#   ungroup() %>%
#   mutate(total = sum(n)) %>%
#   mutate(prop = n / total)

all_interaction_wide %>%
  mutate(Biomass_more_positive_than_energy = ratio_real_Biomass > ratio_real_Energy_use) %>%
  group_by(Same_syndrome, Biomass_more_positive_than_energy) %>%
  tally() %>%
  ungroup() %>%
  group_by(Same_syndrome) %>%
  mutate(total = sum(n))  %>%
  mutate(prop = n / total)

```





## Sample route with different syndromes

Here is an example of a route where biomass shows the "Decoupled trend" syndrome and energy use shows a "coupled trend". For this route, biomass shows a positive slope (ratio of last:first value of 1.22) and energy use a negative slope (ratio of last:first of .9). Individual abundance shows a decline (ratio of last:first of .7). Because the slope for biomass is more positive than for individual abundance (1.22 compared to .7; recall that ratios > 1 are increasing and < 1 are decreasing), we know that community-wide mean body size has increased for this route. Because energy use scales nonlinearly with biomass with an exponent < 1, community-wide mean _per capita metabolic rate_ has not increased as much as mean body size has (but has still increased, because the slope for energy use is less negative than for individual abundance). 


```{r}

different_syndrome_names <- filter(all_interaction_wide, Syndrome_Biomass != Syndrome_Energy_use)

rn <- "bbs_rtrg_6_39"

filter(different_syndrome_names, matssname == rn)

```

```{r, fig.dim = c(5,3)}

dat <- readd(rn, character_only = T, cache=cache)
source(here::here("analyses", "fxns", "is_fxns.R"))

dat_obs <- pull_focal_years(dat)
dat_null <- relabund_null_model(dat_obs)

dat_isd_obs <- simulate_isd_ts(dat_obs, isd_seed = 1989)
dat_isd_null <- simulate_isd_ts(dat_null, isd_seed = 1989)

years <- unique(dat_obs$covariates$year)

begin_years <- years[1:5]
end_years <- years[(length(years) - 4):length(years)]

begin_isd_real <- filter(dat_isd_obs$isd, year %in% begin_years)
end_isd_real <- filter(dat_isd_obs$isd, year %in% end_years )

begin_isd_sim <- filter(dat_isd_null$isd, year %in% begin_years)
end_isd_sim <- filter(dat_isd_null$isd, year %in% end_years)

begin_isd_gmm_real <- add_gmm(begin_isd_real) %>%
  mutate(time = "begin")
end_isd_gmm_real <- add_gmm(end_isd_real) %>%
  mutate(time = "end")

real_gmms <- bind_rows(begin_isd_gmm_real, end_isd_gmm_real) %>%
  mutate(Dynamics = "Observed")

begin_isd_gmm_sim <- add_gmm(begin_isd_sim) %>%
  mutate(time = "begin")
end_isd_gmm_sim <- add_gmm(end_isd_sim)  %>%
  mutate(time = "end")
sim_gmms <- bind_rows(begin_isd_gmm_sim, end_isd_gmm_sim) %>%
  mutate(Dynamics = "Individuals driven")

both_gmms <- bind_rows(real_gmms, sim_gmms) %>%
  mutate(Time = time)
```

This intuition plays out if we calculate the mean body mass and mean metabolic flux off of the ISDs for the first and last 5-year periods. Total abundance has declined (end:begin ratio of .7). Mean metabolic rate and mean biomass have both increased, but the increase in biomass is greater than the increase in metabolic rate (1.63 vs 1.24).

```{r}

both_real_isds <- bind_rows(begin_isd_real, end_isd_real) %>%
  mutate(timeperiod = ifelse(year < 2000, "begin", "end")) %>%
  mutate(metflux = dissBBSsize::estimate_b(mass))


both_real_isds %>%
  group_by(timeperiod) %>%
  summarize(totaln = dplyr::n(),
         mean_body_mass = mean(mass),
         mean_metabolic_flux = mean(metflux))


both_real_isds %>%
  group_by(timeperiod) %>%
  summarize(totaln = dplyr::n(),
         mean_body_mass = mean(mass),
         mean_metabolic_flux = mean(metflux)) %>%
  tidyr::pivot_wider(names_from = timeperiod, values_from = c(totaln, mean_body_mass, mean_metabolic_flux)) %>%
  transmute(totaln_ratio = totaln_end / totaln_begin,
         mean_body_mass_ratio = mean_body_mass_end / mean_body_mass_begin,
         mean_metabolic_flux_ratio = mean_metabolic_flux_end / mean_metabolic_flux_begin)


```

```{r}
isd_plots <- ggplot(filter(both_gmms, exp(mass) > 5, exp(mass) < 5000), aes(exp(mass), density, color = Dynamics, linetype = Time)) + 
  geom_line() + 
  facet_wrap(vars(Dynamics), ncol = 2, scales = "free_y") + 
  theme(legend.position = "bottom") + 
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  ylab("Probability density")+ 
  xlab("Mass (g); note log scale") +
  scale_x_log10()

#isd_plots

```

```{r}


dat_dynamics <- filter(all_sims, matssname == rn)

dat_glm <- glm(total_biomass ~ timeperiod * source, data = dat_dynamics, family= Gamma(link = "log"))

dat_lm <- lm(total_biomass ~ timeperiod * source, data = dat_dynamics)


dat_dynamics <- dat_dynamics %>%
  mutate(fitted = predict(dat_glm, type = "response")) %>%
  mutate(Dynamics = ifelse(source == "sim", "Abundance driven", "Observed")) %>%
  mutate(time = "begin") 

timeseries_plot <- ggplot(dat_dynamics, aes(timeperiod, total_biomass, color = Dynamics)) +
  geom_point() + 
  geom_line() + 
  geom_line(aes(y = fitted)) + 
  # geom_rect(aes(xmin = years[1], xmax = years[5], ymin = min(total_biomass), ymax = max(total_biomass), time = "begin"), alpha = 0, color = "black")+ 
  # geom_rect(data=  mutate(dat_dynamics, time = "end"), aes(xmin = years[length(years) - 4], xmax = years[length(years)], ymin = min(total_biomass), ymax = max(total_biomass),linetype = time), alpha = 0, color = "black")+
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Biomass")


dat_glm_e <- glm(total_energy ~ timeperiod * source, data = dat_dynamics, family= Gamma(link = "log"))


dat_dynamics_e <- dat_dynamics %>%
  mutate(fitted = predict(dat_glm_e, type = "response")) %>%
  mutate(Dynamics = ifelse(source == "sim", "Abundance driven", "Observed")) %>%
  mutate(time = "begin") 

timeseries_plot_e <- ggplot(dat_dynamics_e, aes(timeperiod, total_energy, color = Dynamics)) +
  geom_point() + 
  geom_line() + 
  geom_line(aes(y = fitted)) + 
  # geom_rect(aes(xmin = years[1], xmax = years[5], ymin = min(total_biomass), ymax = max(total_biomass), time = "begin"), alpha = 0, color = "black")+ 
  # geom_rect(data=  mutate(dat_dynamics, time = "end"), aes(xmin = years[length(years) - 4], xmax = years[length(years)], ymin = min(total_biomass), ymax = max(total_biomass),linetype = time), alpha = 0, color = "black")+
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Energy use")



#timeseries_plot
```

```{r, fig.dim = c(6,9)}
two_plots <- multi_panel_figure(columns = 1 ,height = c(2,2,2.25), unit = "in") %>%
  fill_panel(timeseries_plot) %>%
  fill_panel(timeseries_plot_e) %>%
  fill_panel(isd_plots)

two_plots
```



## Sample route with the same syndrome

Contrast this with a route where biomass and energy use show the same "syndrome" and the same general dynamics. For this route, both biomass and energy use show the "decoupled trends" syndrome. Abundance has declined, but both biomass and energy use have increased.

My intuition is that when and why changes in energy use and biomass are similar enough to show the same or different broad types of dynamics come down to the nuances of where in the size spectrum change occurs. If (as is the most common scenario in this dataset) mean individual body size increases, mean metabolic flux will also increase but to a lesser degree due to the sub-1 scaling relationship. Depending on 

```{r}

same_syndrome_names <- filter(all_interaction_wide, Syndrome_Biomass == Syndrome_Energy_use)
# (filter(same_syndrome_names, modelcomplexity_Biomass == 4, ratio_real_Biomass > 1, ratio_sim_Biomass < 1))
rn <- "bbs_rtrg_44_17"

filter(same_syndrome_names, matssname == rn)

```

```{r, fig.dim = c(5,3)}

dat <- readd(rn, character_only = T, cache=cache)
source(here::here("analyses", "fxns", "is_fxns.R"))

dat_obs <- pull_focal_years(dat)
dat_null <- relabund_null_model(dat_obs)

dat_isd_obs <- simulate_isd_ts(dat_obs, isd_seed = 1989)
dat_isd_null <- simulate_isd_ts(dat_null, isd_seed = 1989)

years <- unique(dat_obs$covariates$year)

begin_years <- years[1:5]
end_years <- years[(length(years) - 4):length(years)]

begin_isd_real <- filter(dat_isd_obs$isd, year %in% begin_years)
end_isd_real <- filter(dat_isd_obs$isd, year %in% end_years )

begin_isd_sim <- filter(dat_isd_null$isd, year %in% begin_years)
end_isd_sim <- filter(dat_isd_null$isd, year %in% end_years)

begin_isd_gmm_real <- add_gmm(begin_isd_real) %>%
  mutate(time = "begin")
end_isd_gmm_real <- add_gmm(end_isd_real) %>%
  mutate(time = "end")

real_gmms <- bind_rows(begin_isd_gmm_real, end_isd_gmm_real) %>%
  mutate(Dynamics = "Observed")

begin_isd_gmm_sim <- add_gmm(begin_isd_sim) %>%
  mutate(time = "begin")
end_isd_gmm_sim <- add_gmm(end_isd_sim)  %>%
  mutate(time = "end")
sim_gmms <- bind_rows(begin_isd_gmm_sim, end_isd_gmm_sim) %>%
  mutate(Dynamics = "Individuals driven")

both_gmms <- bind_rows(real_gmms, sim_gmms) %>%
  mutate(Time = time)
```

```{r}

both_real_isds <- bind_rows(begin_isd_real, end_isd_real) %>%
  mutate(timeperiod = ifelse(year < 2000, "begin", "end")) %>%
  mutate(metflux = dissBBSsize::estimate_b(mass))


both_real_isds %>%
  group_by(timeperiod) %>%
  summarize(totaln = dplyr::n(),
         mean_body_mass = mean(mass),
         mean_metabolic_flux = mean(metflux))


both_real_isds %>%
  group_by(timeperiod) %>%
  summarize(totaln = dplyr::n(),
         mean_body_mass = mean(mass),
         mean_metabolic_flux = mean(metflux)) %>%
  tidyr::pivot_wider(names_from = timeperiod, values_from = c(totaln, mean_body_mass, mean_metabolic_flux)) %>%
  transmute(totaln_ratio = totaln_end / totaln_begin,
         mean_body_mass_ratio = mean_body_mass_end / mean_body_mass_begin,
         mean_metabolic_flux_ratio = mean_metabolic_flux_end / mean_metabolic_flux_begin)


```

```{r}
isd_plots <- ggplot(filter(both_gmms, exp(mass) > 5, exp(mass) < 5000), aes(exp(mass), density, color = Dynamics, linetype = Time)) + 
  geom_line() + 
  facet_wrap(vars(Dynamics), ncol = 2, scales = "free_y") + 
  theme(legend.position = "bottom") + 
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  ylab("Probability density")+ 
  xlab("Mass (g); note log scale") +
  scale_x_log10()

#isd_plots

```

```{r}


dat_dynamics <- filter(all_sims, matssname == rn)

dat_glm <- glm(total_biomass ~ timeperiod * source, data = dat_dynamics, family= Gamma(link = "log"))

dat_lm <- lm(total_biomass ~ timeperiod * source, data = dat_dynamics)


dat_dynamics <- dat_dynamics %>%
  mutate(fitted = predict(dat_glm, type = "response")) %>%
  mutate(Dynamics = ifelse(source == "sim", "Abundance driven", "Observed")) %>%
  mutate(time = "begin") 

timeseries_plot <- ggplot(dat_dynamics, aes(timeperiod, total_biomass, color = Dynamics)) +
  geom_point() + 
  geom_line() + 
  geom_line(aes(y = fitted)) + 
  # geom_rect(aes(xmin = years[1], xmax = years[5], ymin = min(total_biomass), ymax = max(total_biomass), time = "begin"), alpha = 0, color = "black")+ 
  # geom_rect(data=  mutate(dat_dynamics, time = "end"), aes(xmin = years[length(years) - 4], xmax = years[length(years)], ymin = min(total_biomass), ymax = max(total_biomass),linetype = time), alpha = 0, color = "black")+
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Biomass")


dat_glm_e <- glm(total_energy ~ timeperiod * source, data = dat_dynamics, family= Gamma(link = "log"))


dat_dynamics_e <- dat_dynamics %>%
  mutate(fitted = predict(dat_glm_e, type = "response")) %>%
  mutate(Dynamics = ifelse(source == "sim", "Abundance driven", "Observed")) %>%
  mutate(time = "begin") 

timeseries_plot_e <- ggplot(dat_dynamics_e, aes(timeperiod, total_energy, color = Dynamics)) +
  geom_point() + 
  geom_line() + 
  geom_line(aes(y = fitted)) + 
  # geom_rect(aes(xmin = years[1], xmax = years[5], ymin = min(total_biomass), ymax = max(total_biomass), time = "begin"), alpha = 0, color = "black")+ 
  # geom_rect(data=  mutate(dat_dynamics, time = "end"), aes(xmin = years[length(years) - 4], xmax = years[length(years)], ymin = min(total_biomass), ymax = max(total_biomass),linetype = time), alpha = 0, color = "black")+
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Energy use")



#timeseries_plot
```

```{r, fig.dim = c(6,9)}
two_plots <- multi_panel_figure(columns = 1 ,height = c(2,2,2.25), unit = "in") %>%
  fill_panel(timeseries_plot) %>%
  fill_panel(timeseries_plot_e) %>%
  fill_panel(isd_plots)

two_plots
```
