---
output:
  # word_document:
  #   reference_docx: default_gdoc.docx
  #   df_print: kable
  # html_document:
  #   df_print: kable
  pdf_document: 
    df_print: kable
csl: ecology.csl
bibliography: refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = T, results = T, message = F, warning = F, eval = T, fig.dim = c(4,4))
library(dplyr)
library(drake)
library(multipanelfigure)
library(ggplot2)
theme_set(theme_bw())

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analyses", "caches", "all_hasty_toy.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")


```

# Abundance-driven vs. observed change

```{r, fig.dim = c(5,3)}

loadd(all_sims, cache=cache)
rn <- "bbs_rtrg_28_4"
dat <- readd(rn, character_only = T, cache=cache)
source(here::here("analyses", "fxns", "hasty_fxns.R"))

dat_sliced <- slice_dat(dat)
dat_resp <- resp_dat(dat_sliced)

dat_isd <- simulate_isd_ts(dat_sliced, isd_seed = 1989)
dat_resp_isd <- simulate_isd_ts(dat_resp, isd_seed = 1989)

years <- unique(dat_sliced$covariates$year)

begin_years <- years[1:5]
end_years <- years[(length(years) - 4):length(years)]

begin_isd_real <- filter(dat_isd$isd, year %in% begin_years)
end_isd_real <- filter(dat_isd$isd, year %in% end_years )

begin_isd_sim <- filter(dat_resp_isd$isd, year %in% begin_years)
end_isd_sim <- filter(dat_resp_isd$isd, year %in% end_years)

begin_isd_gmm_real <- add_gmm(begin_isd_real) %>%
  mutate(time = "begin")
end_isd_gmm_real <- add_gmm(end_isd_real) %>%
  mutate(time = "end")

real_gmms <- bind_rows(begin_isd_gmm_real, end_isd_gmm_real) %>%
  mutate(simtype = "Observed")

begin_isd_gmm_sim <- add_gmm(begin_isd_sim) %>%
  mutate(time = "begin")
end_isd_gmm_sim <- add_gmm(end_isd_sim)  %>%
  mutate(time = "end")
sim_gmms <- bind_rows(begin_isd_gmm_sim, end_isd_gmm_sim) %>%
  mutate(simtype = "Abundance driven")

both_gmms <- bind_rows(real_gmms, sim_gmms)
```

```{r}
isd_plots <- ggplot(filter(both_gmms, exp(mass) > 5, exp(mass) < 5000), aes(exp(mass), density, color = simtype, linetype = time)) + 
  geom_line() + 
  facet_wrap(vars(simtype), ncol = 2, scales = "free_y") + 
  theme(legend.position = "bottom") + 
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  ylab("Probability density")+ 
  xlab("Mass (g); note log scale") +
  scale_x_log10()

#isd_plots

```

```{r}


dat_dynamics <- filter(all_sims, matssname == rn)

dat_glm <- glm(total_biomass ~ timeperiod * source, data = dat_dynamics, family= Gamma(link = "log"))

dat_lm <- lm(total_biomass ~ timeperiod * source, data = dat_dynamics)


dat_dynamics <- dat_dynamics %>%
  mutate(fitted = predict(dat_glm, type = "response")) %>%
  mutate(simtype = ifelse(source == "sim", "Abundance driven", "Observed")) %>%
  mutate(time = "begin") 

timeseries_plot <- ggplot(dat_dynamics, aes(timeperiod, total_biomass, color = simtype)) +
  geom_point() + 
  geom_line() + 
  geom_line(aes(y = fitted)) + 
  # geom_rect(aes(xmin = years[1], xmax = years[5], ymin = min(total_biomass), ymax = max(total_biomass), time = "begin"), alpha = 0, color = "black")+ 
  # geom_rect(data=  mutate(dat_dynamics, time = "end"), aes(xmin = years[length(years) - 4], xmax = years[length(years)], ymin = min(total_biomass), ymax = max(total_biomass),linetype = time), alpha = 0, color = "black")+
  scale_color_viridis_d(option = "cividis",begin = .2,  end = .8) + 
  theme(legend.position = "none") +
  xlab("Year") +
  ylab("Total biomass")

#timeseries_plot
```

```{r, fig.dim = c(6,4.25)}
two_plots <- multi_panel_figure(columns = 1 ,height = c(2,2.25), unit = "in") %>%
  fill_panel(timeseries_plot) %>%
  fill_panel(isd_plots)

two_plots
```

#### Figure 1. 

Illustration of abundance-driven (null model) dynamics as compared to observed dynamics (A), and the underlying dynamics of the ISD (B) for a sample route (`r dat_isd$metadata$location$routename`, `r dat_isd$metadata$location$regionname`). **A. Dynamics of total biomass.** The orange points show the true values for total biomass in each year, and the purple points show the values for total biomass simulated from a null model that incorporates change in total abundance, but assumes no change in the size structure, over time. The smooth lines show the predicted values from a Gamma (log-link) linear model of the form `total_biomass ~ year * simtype`. <!-- Using gaussian here because it is easier to interpret directly and compute r2. results are the same if you use Gamma --> For this route, change in the individual size distribution has decoupled the dynamics of biomass from those that would occur due only to changes in abundance. The slope for abundance-driven dynamics is significantly more negative than for the observed dynamics (interaction term p = 0.0013), which do not have a slope significantly different from zero (slope term p = .27). **B. Underlying changes in the ISD.** The individual size distributions for the first 5 years (solid lines) and last 5 years (dashed lines) of the timeseries. The x-axis is body size (as mass in grams; note log scale) and the y-axis is probability density from a Gaussian mixture model fit to a vector of simulated individual masses for all individuals observed in the years in questions, standardized to sum to 1. For the abundance-driven (purple) scenario, individuals' species identities (which determine their body size estimates) are re-assigned at random weighte by each species' mean relative abundance throughout the timeseries, resulting in a consistent individual size distribution over time. For the observed (orange) scenario, individuals' body sizes are estimated based on their actual species identities. For this route, species composition has shifted over time, resulting in distinct ISDs for the "begin" and "end" time periods. Specifically, the "end" ISD has peaks at larger body sizes (ca. 90g and 500g) not present in the "begin" ISD. This redistribution of density towards larger body sizes results in an overall increase in body size community wise, which partially offsets declines in total biomass from those expected given change in abundance alone. 

\newpage


```{r load and process biomass results}

loadd(all_preds, cache=cache)
loadd(all_aics, cache=cache)

winning_aics <- all_aics %>%
  filter(model_family == "Gamma") %>%
  group_by(matssname, model_family) %>%
  mutate(minAICc = min(model_AICc)) %>%
  mutate(deltaAICc = minAICc - model_AICc)  %>%
  mutate(exp_deltas = exp(.5 * deltaAICc)) %>%
  mutate(denom = sum(exp_deltas)) %>%
  mutate(aicc_wt = exp_deltas/ denom) %>%
  arrange(desc(aicc_wt)) %>%
  mutate(rank =row_number()) %>%
  ungroup() %>%
  arrange(matssname, model_family, rank)

delta_2 <- winning_aics %>% filter(deltaAICc > -2) %>%
  filter(model_family == "Gamma") %>%
  group_by(matssname, model_family) %>%
  arrange(modelcomplexity) %>%
  mutate(rank = row_number()) %>%
  filter(rank == 1) %>%
  ungroup()

winning_fits <- delta_2 %>%
  left_join(all_preds) 

slope_winning_fits <- winning_fits %>%
  filter(modelcomplexity > 1)

interaction_winning_fits <- winning_fits %>%
  filter(modelcomplexity > 2)


biomass_slopes_long <- slope_winning_fits %>%
  select(matssname, ratio_sim, ratio_real) %>%
  tidyr::pivot_longer(-matssname, names_to = "source", values_to = "biomass_ratio") %>%
  mutate(simtype = ifelse(source == "ratio_sim", "Abundance driven", "Observed"))

```


```{r load and process energy results}

loadd(all_preds_energy, cache=cache)
loadd(all_aics_energy, cache=cache)

winning_aics_energy <- all_aics_energy %>%
  filter(model_family == "Gamma") %>%
  group_by(matssname, model_family) %>%
  mutate(minAICc = min(model_AICc)) %>%
  mutate(deltaAICc = minAICc - model_AICc)  %>%
  mutate(exp_deltas = exp(.5 * deltaAICc)) %>%
  mutate(denom = sum(exp_deltas)) %>%
  mutate(aicc_wt = exp_deltas/ denom) %>%
  arrange(desc(aicc_wt)) %>%
  mutate(rank =row_number()) %>%
  ungroup() %>%
  arrange(matssname, model_family, rank)

delta_2_energy <- winning_aics_energy %>% filter(deltaAICc > -2) %>%
  filter(model_family == "Gamma") %>%
  group_by(matssname, model_family) %>%
  arrange(modelcomplexity) %>%
  mutate(rank = row_number()) %>%
  filter(rank == 1) %>%
  ungroup()


winning_fits_energy <- delta_2_energy %>%
  left_join(all_preds_energy)

slope_winning_fits_energy <- winning_fits_energy %>%
  filter(modelcomplexity > 1)

interaction_winning_fits_energy <- winning_fits_energy %>%
  filter(modelcomplexity > 2)


energy_slopes_long <- slope_winning_fits_energy %>%
  select(matssname, ratio_sim, ratio_real) %>%
  tidyr::pivot_longer(-matssname, names_to = "source", values_to = "energy_ratio") %>%
  mutate(simtype = ifelse(source == "ratio_sim", "Abundance driven", "Observed"))
```


\newpage

# Slopes

```{r, fig.dim = c(4,4)}

biomass_slope_histogram <- ggplot(biomass_slopes_long, aes(biomass_ratio, color = simtype, fill = simtype)) + geom_histogram(boundary = 1) + scale_x_log10() + geom_vline(xintercept = 1) + facet_wrap(vars(simtype), scales = "free_y") + scale_color_viridis_d(option = "cividis", begin = .2, end = .8) + scale_fill_viridis_d(option = "cividis", begin = .2, end = .8) +
  theme(legend.position = "none") + ggtitle("Total biomass") + xlab("")


energy_slope_histogram <- ggplot(energy_slopes_long, aes(energy_ratio, color = simtype, fill = simtype)) + geom_histogram(boundary = 1) + scale_x_log10() + geom_vline(xintercept = 1) + facet_wrap(vars(simtype), scales = "free_y") + scale_color_viridis_d(option = "cividis", begin = .2, end = .8) + scale_fill_viridis_d(option = "cividis", begin = .2, end = .8) +
  theme(legend.position = "bottom") + ggtitle("Total metabolic flux") + xlab("Ratio of last fitted value to first fitted value")

slope_histograms <- multi_panel_figure(columns = 1, height = c(1.6, 2), unit = "in") %>%
  fill_panel(biomass_slope_histogram) %>%
  fill_panel(energy_slope_histogram)
slope_histograms
```

#### Figure 2. Long-term trends in total biomass and energy use

Histograms showing the direction and magnitude of change over time for the abundance-driven (left) and observed (right) changes in biomass (A) and energy use (B), for communities with a significant slope and/or interaction term (for biomass, 500/739 routes; for energy use, 506/739 routes). Change is summarized as the ratio of the fitted value for the last year in the time series to the fitted value for the first year in the timeseries from the best-fitting model for that community. Values greater than 1 (vertical black line) indicate increases in total energy or biomass over time, and less than 1 indicate decreases. The abundance-driven dynamics (left) reflect the trends fit for the null model, while the observed dynamics (right) reflect trends incorporating both change in total abundance and change in the size structure over time. For communities with no significant interaction term in the best-fitting model, the "abundance-driven" and "observed" ratios will be the same; interaction terms will result in different ratios for "abundance-driven" and "observed" dynamics. 

Abundance-driven dynamics are dominated by decreases. For biomass, observed dynamics are balanced evenly between increases and decreases - indicating that changes in the size structure produce qualitatively different long-term trends for biomass than would be expected given abundance changes alone. However, trends for energy use (which scales nonlinearly with biomass) are dominated by decreases, more closely mirroring the trends expected given changes in abundance. 

\newpage

# Model outcomes 

## Biomass

Overall proportion of routes with winning models:

```{r, results =T}

winning_fits %>% group_by(model_family, model_formula) %>% tally() %>%
  mutate(prop = n / sum(n))

```

Of models with slope term, the proportion for which abundance and biomass are increasing:

```{r}

slope_winning_fits %>%
  mutate(abundance_increasing = ratio_sim > 1) %>%
  group_by(abundance_increasing) %>%
  tally() %>%
  mutate(prop = n / sum(n))


slope_winning_fits %>%
  mutate(biomass_increasing = ratio_real > 1) %>%
  group_by(biomass_increasing) %>%
  tally() %>%
  mutate(prop = n / sum(n))
```


<!-- Restricted to models with an *interaction*: -->

<!-- ```{r} -->
<!-- interaction_winning_fits %>% -->
<!--   mutate(abundance_increasing = ratio_sim > 1) %>% -->
<!--   group_by(abundance_increasing) %>% -->
<!--   tally() %>% -->
<!--   mutate(prop = n / sum(n)) -->


<!-- interaction_winning_fits %>% -->
<!--   mutate(biomass_increasing = ratio_real > 1) %>% -->
<!--   group_by(biomass_increasing) %>% -->
<!--   tally() %>% -->
<!--   mutate(prop = n / sum(n)) -->

<!-- ``` -->

## Metabolic flux

Overall proportion of routes with winning models:

```{r, results =T}

winning_fits_energy %>% group_by(model_family, model_formula) %>% tally() %>%
  mutate(prop = n / sum(n))

```

Of models with slope term, the proportion for which abundance and biomass are increasing:

```{r}

slope_winning_fits_energy %>%
  mutate(abundance_increasing = ratio_sim > 1) %>%
  group_by(abundance_increasing) %>%
  tally() %>%
  mutate(prop = n / sum(n))

slope_winning_fits_energy %>%
  mutate(energy_increasing = ratio_real > 1) %>%
  group_by(energy_increasing) %>%
  tally() %>%
  mutate(prop = n / sum(n))
```


<!-- Restricted to models with an *interaction*: -->

<!-- ```{r} -->
<!-- interaction_winning_fits_energy %>% -->
<!--   mutate(abundance_increasing = ratio_sim > 1) %>% -->
<!--   group_by(abundance_increasing) %>% -->
<!--   tally() %>% -->
<!--   mutate(prop = n / sum(n)) -->


<!-- interaction_winning_fits_energy %>% -->
<!--   mutate(energy_increasing = ratio_real > 1) %>% -->
<!--   group_by(energy_increasing) %>% -->
<!--   tally() %>% -->
<!--   mutate(prop = n / sum(n)) -->

<!-- ``` -->

\newpage

# Visualizing decoupling

```{r}

interaction_all_fits <- all_preds %>% 
  # get just the change ratios from the interaction models - regardless of whether the interaction model is the best-fitting model. this is because we want an estimate of the slope, we'll color by model fit in a minute
  group_by_all() %>% 
  filter(grepl( "* source",model_formula), model_family == "Gamma") %>%
  mutate(plus = substr(model_formula, 12,12) == "+") %>% filter(!plus) %>%
  ungroup() %>%
  # now join to *winning* formulas to get the best fitting formula
  left_join(select(winning_fits, matssname, model_formula, modelcomplexity) %>%
              rename(winning_formula = model_formula)) %>%
  select(winning_formula, ratio_sim, ratio_real, matssname, modelcomplexity) %>%
  mutate(currency = "Biomass")


interaction_all_fits_energy <- all_preds_energy  %>% 
  # get just the change ratios from the interaction models - regardless of whether the interaction model is the best-fitting model. this is because we want an estimate of the slope, we'll color by model fit in a minute
  group_by_all() %>% 
  filter(grepl( "* source",model_formula), model_family == "Gamma") %>%
  mutate(plus = substr(model_formula, 12,12) == "+") %>% filter(!plus) %>%
  ungroup() %>%
  # now join to *winning* formulas to get the best fitting formula
  left_join(select(winning_fits_energy, matssname, model_formula, modelcomplexity) %>%
              rename(winning_formula = model_formula)) %>%
  select(winning_formula, ratio_sim, ratio_real, matssname, modelcomplexity) %>%
  mutate(currency = "Metabolic flux")


all_interaction <- bind_rows(interaction_all_fits, interaction_all_fits_energy) %>%
  mutate(`Best fitting model` = ifelse(modelcomplexity == 1, "No trend", ifelse(modelcomplexity == 2, "Coupled trend", "Decoupled trends")))


```


```{r, fig.dim= c(5.5,3)}

decoupling_plot <- ggplot(all_interaction, aes(ratio_sim, ratio_real, color = `Best fitting model`)) + 
  geom_point(alpha = .5) + 
  #geom_point(data = filter(all_interaction, winning_formula != "1"), aes(color = winning_formula), alpha = .3) + 
  scale_y_log10() + 
  scale_x_log10() + 
  scale_color_viridis_d(option = "mako", end = .8, direction = -1) + 
  theme(legend.position = "bottom") + 
  geom_abline(intercept = 0, slope = 1) + 
  geom_vline(xintercept = 1) + 
  geom_hline(yintercept = 1) +
  xlab("Abundance-driven change \n Ratio of last fitted value to first fitted value") + 
  ylab("Observed change \n Ratio of last fitted value to first fitted value") +
  facet_wrap(vars(currency), ncol = 2)


decoupling_plot
```


# How the ISD contributes to decoupling

```{r}

loadd(all_isd_compares, cache =cache)

winning_fits <- winning_fits %>% 
  left_join(all_isd_compares) %>%
  mutate(`Best fitting model` = ifelse(modelcomplexity == 1, "No trend", ifelse(modelcomplexity == 2, "Coupled trend", "Decoupled trends")))


isd_mean_change_plot <-  ggplot(winning_fits, aes(real_mass_ratio, color = `Best fitting model`, fill= `Best fitting model`)) + 
  geom_histogram() + 
  facet_wrap(vars(`Best fitting model`), scales = 'free_y', ncol = 1) +
  scale_color_viridis_d(option = "mako", end = .8, direction = -1) +     
  scale_fill_viridis_d(option = "mako", end = .8, direction = -1) +
  theme(legend.position = "none") +
  xlab("Ratio of mean mass\n(last five years / first five years)") +
  ylab("Count") +
  ggtitle("Change in mean body size")

#isd_mean_change_plot


```

```{r}


isd_overlap_plot <- ggplot(winning_fits, aes(real_overlap,color = `Best fitting model`, fill= `Best fitting model`)) + 
  geom_histogram() + 
  facet_wrap(vars(`Best fitting model`), scales = 'free_y', ncol = 1) +
  scale_color_viridis_d(option = "mako", end = .8, direction = -1) +  scale_fill_viridis_d(option = "mako", end = .8, direction = -1) +
  theme(legend.position = "none") +
  xlab("ISD overlap\n(last five years vs. first five years)") +
  ylab("Count") +
  ggtitle("Overall change in ISD")
 # isd_overlap_plot
# 
# isd_glm <- glm(real_overlap ~ model_formula, family = "binomial", data = winning_fits)
# #
# # AIC(isd_glm)
# # AIC(glm(real_overlap ~ 1, family = "binomial", data = winning_fits))
# 
# glm_cor <- cor(predict(isd_glm, type = "response"), isd_glm$model[,1]) ^ 2
```

```{r, fig.dim = c(5.5, 4)}

isd_change_plots <- multi_panel_figure(columns = 2, rows = 1) %>%
  fill_panel(isd_mean_change_plot) %>%
  fill_panel(isd_overlap_plot)
isd_change_plots
```

R2 of binomial GLM `overlap ~ model_formula` (which does not beat a `overlap ~ 1` via AIC, FYI)

```{r}
#glm_cor
```

<!-- # Goodness of fit of models -->

<!-- ```{r} -->

<!-- ggplot(winning_fits, aes(model_cor_squared)) + geom_histogram() + facet_wrap(vars(model_formula), scales = "free_y") -->

<!-- ggplot(winning_fits_energy, aes(model_cor_squared)) + geom_histogram() + facet_wrap(vars(model_formula), scales = "free_y") -->

<!-- ``` -->


```{r}

winning_fits <- winning_fits %>%
  mutate(categorical_fit = `Best fitting model`) %>%
  mutate(categorical_fit = as.factor(categorical_fit))


```


\newpage
# References
