---
output:
  word_document:
    reference_docx: ecology_doc.docx
    df_print: kable
  # html_document:
  #   df_print: kable
  pdf_document: 
    df_print: kable
csl: ecology.csl
bibliography: refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = T, results = T, message = F, warning = F, eval = T, fig.dim = c(4,4))
library(dplyr)
library(drake)
library(multipanelfigure)
library(ggplot2)
theme_set(theme_bw())

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analyses", "caches", "all.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)
cache$del(key = "lock", namespace = "session")


```


```{r load results and filter to desired temporal coverage}

desired_routes <- read.csv(here::here("analyses", "supporting_data", "perfect_coverage_1988_2018.csv"), stringsAsFactors = F)

loadd(all_sims, cache=cache)
rn <- "bbs_rtrg_28_4"

dat_dynamics <- filter(all_sims, matssname == rn)

all_sims <- all_sims %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_preds_b, cache=cache)
all_preds_b <- all_preds_b %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_aics_b, cache=cache)
all_aics_b <- all_aics_b %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_preds_e, cache=cache)
all_preds_e <- all_preds_e %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_aics_e, cache=cache)
all_aics_e <- all_aics_e %>%
  filter(matssname %in% desired_routes$matssname)

loadd(all_cs_compares, cache =cache)
all_cs_compares <- all_cs_compares %>%
  filter(matssname %in% desired_routes$matssname)

```


```{r process biomass results}

# Select the best-fitting model for each route using AICc (simplest model with an AICc score within two units of the minimum)
# all_aics contains the AIC and AICc scores for every model fit to each route.

winning_aics <- all_aics_b %>%
  filter(model_family == "Gamma") %>% # I fit both Gamma and Gaussian models. The Gaussian are easier to interpret (slope terms can be read out directly) but are less strictly appropriate for strictly-positive biomass/energy use data.
  group_by(matssname, model_family) %>%
  mutate(minAICc = min(model_AICc)) %>% # identify best AICc
  mutate(deltaAICc = minAICc - model_AICc)  %>% # deltaAICc is the difference between each model's AICc and the best AICc for that pool
  filter(deltaAICc > -2) %>% # use 2 as a rule of thumb cutoff 
  arrange(matssname, modelcomplexity) %>% # rank by model complexity. 1 is simplest, 4 is most complex.
  mutate(rank = row_number()) %>% 
  filter(rank == 1) %>% # select the simplest model with deltaAICc within 2 of the best
  ungroup()


# Get predicted changes for winning models
# all_preds contains predicted trends for all models; the left_join effectively filters to the models in winning_aics.
winning_fits <- winning_aics %>%
  left_join(all_preds_b) 


# Get changes for winning models *only for routes for which there is some temporal trend*; i.e. the winning model is not the intercept-only model.
slope_winning_fits <- winning_fits %>%
  filter(modelcomplexity > 1)

# Pivot longer for plotting and summarizing
biomass_slopes_long <- slope_winning_fits %>%
  select(matssname, ratio_sim, ratio_real) %>%
  tidyr::pivot_longer(-matssname, names_to = "source", values_to = "biomass_ratio") %>%
  mutate(Dynamics = ifelse(source == "ratio_sim", "Abundance driven", "Observed"))

```


```{r process energy results}


winning_aics_energy <- all_aics_e %>%
  filter(model_family == "Gamma") %>% # I fit both Gamma and Gaussian models. The Gaussian are easier to interpret (slope terms can be read out directly) but are less strictly appropriate for strictly-positive biomass/energy use data.
  group_by(matssname, model_family) %>%
  mutate(minAICc = min(model_AICc)) %>% # identify best AICc
  mutate(deltaAICc = minAICc - model_AICc)  %>% # deltaAICc is the difference between each model's AICc and the best AICc for that pool
  filter(deltaAICc > -2) %>% # use 2 as a rule of thumb cutoff 
  arrange(matssname, modelcomplexity) %>% # rank by model complexity. 1 is simplest, 4 is most complex.
  mutate(rank = row_number()) %>% 
  filter(rank == 1) %>% # select the simplest model with deltaAICc within 2 of the best
  ungroup()


# Get predicted changes for winning models
# all_preds contains predicted trends for all models; the left_join effectively filters to the models in winning_aics.
winning_fits_energy <- winning_aics_energy %>%
  left_join(all_preds_e) 


# Get changes for winning models *only for routes for which there is some temporal trend*; i.e. the winning model is not the intercept-only model.
slope_winning_fits_energy <- winning_fits_energy %>%
  filter(modelcomplexity > 1)

# Pivot longer for plotting and summarizing
energy_slopes_long <- slope_winning_fits_energy %>%
  select(matssname, ratio_sim, ratio_real) %>%
  tidyr::pivot_longer(-matssname, names_to = "source", values_to = "energy_ratio") %>%
  mutate(Dynamics = ifelse(source == "ratio_sim", "Abundance driven", "Observed"))


# Combine energy and biomass results into one dataframe
winning_fits_both <- winning_fits %>%
  mutate(currency = "biomass") %>%
  bind_rows(mutate(winning_fits_energy, currency = "energy"))


```


\newpage

# Appendix S1 Figure S1. 

```{r, fig.dim = c(4,4)}

biomass_slope_histogram <- ggplot(biomass_slopes_long, aes(biomass_ratio, color = Dynamics, fill = Dynamics)) + geom_histogram(boundary = 1) + scale_x_log10() + geom_vline(xintercept = 1) + facet_wrap(vars(Dynamics), scales = "free_y") + scale_color_viridis_d(option = "cividis", begin = .2, end = .8) + scale_fill_viridis_d(option = "cividis", begin = .2, end = .8) +
  theme(legend.position = "none") + ggtitle("Total biomass") + xlab("")


energy_slope_histogram <- ggplot(energy_slopes_long, aes(energy_ratio, color = Dynamics, fill = Dynamics)) + geom_histogram(boundary = 1) + scale_x_log10() + geom_vline(xintercept = 1) + facet_wrap(vars(Dynamics), scales = "free_y") + scale_color_viridis_d(option = "cividis", begin = .2, end = .8) + scale_fill_viridis_d(option = "cividis", begin = .2, end = .8) +
  theme(legend.position = "bottom") + ggtitle("Total metabolic flux") + xlab("Ratio of last fitted value to first fitted value")

slope_histograms <- multi_panel_figure(columns = 1, height = c(1.6, 2), unit = "in") %>%
  fill_panel(biomass_slope_histogram) %>%
  fill_panel(energy_slope_histogram)
slope_histograms
```

#### Appendix S1 Figure S1. Long-term trends in total biomass and energy use.
Histograms showing the direction and magnitude of long-term trends for the abundance-driven (null-model; left) and observed (right) changes in biomass (A) and energy use (B), for communities with a significant slope and/or interaction term (for biomass, 141/199 routes; for energy use, 137/199 routes; Table 1). Change is summarized as the ratio of the fitted value for the last year in the time series to the fitted value for the first year in the timeseries from the best-fitting model for that community. Values greater than 1 (vertical black line) indicate increases in total energy or biomass over time, and less than 1 indicate decreases. The abundance-driven dynamics (left) reflect the trends fit for the null model, while the observed dynamics (right) reflect trends incorporating both change in total abundance and change in the size structure over time. For communities best-described by syndromes of “coupled trends” or “no directional change”, the “abundance-driven” and “observed” ratios will be the same; for communities with “decoupled trends”, there will be different ratios for or “abundance-driven” and “observed” dynamics.

Among routes with temporal trends (“coupled trends” or “decoupled trends”), there are qualitatively different continental-wide patterns in abundance-driven and observed dynamics for total biomass and total energy use. 76% of trends in abundance-driven (null model) dynamics for energy use are decreasing, and 72% for biomass (Table 2). For biomass, observed dynamics are balanced evenly between increases (50% of routes) and decreases (50%) - indicating that changes in the size structure produce qualitatively different long-term trends for biomass than would be expected given abundance changes alone. However, trends for energy use (which scales nonlinearly with biomass) are dominated by decreases (69% of routes), more closely mirroring the trends expected given changes in individual abundance alone.


\newpage

# Tables

### Appendix S1 Table S1.

```{r model outcomes table, results = T}

model_outcomes_table <- winning_fits_both %>%
  select(matssname, model_family, model_formula, currency)  %>%
  group_by(currency, model_formula) %>%
  summarize(n = dplyr::n()) %>%
  ungroup() %>%
  group_by(currency) %>%
  mutate(nmods = sum(n)) %>%
  mutate(prop = n / nmods) 

model_outcomes_table_pretty <- model_outcomes_table %>%
  mutate(currency = ifelse(currency == "biomass", "Total biomass", "Total metabolic flux")) %>%
  mutate(prop_rounded = round(prop, digits = 2),
         form_pretty = ifelse(grepl("1", model_formula), "Intercept-only", ifelse(grepl("source", model_formula), "Decoupled trend", "Trend, not decoupled"))) %>%
  select(currency, form_pretty, n, prop_rounded) %>%
  rename(`Selected model` = form_pretty,
         `Number of routes` = n,
         `Proportion of routes` = prop_rounded,
         Currency = currency) 

model_outcomes_table_pretty

```
#### Table 1. Selected models.

Table of the number and proportion of routes whose dynamics for total biomass and total energy use are best-fit by: a model with no temporal trend (intercept-only model, `response ~ 1`); a model with a temporal trend, but no difference in trend between observed and abundance-driven dynamics (`response ~ timeperiod`); or a model with decoupled temporal trends for observed and abundance-driven dynamics (`response ~ timeperiod * dynamics`, where `dynamics` is either observed or abundance-driven/null model). 

For 31-32% of routes, models with trends do not outperform simple intercept-only models. For the remaining routes, in most instances, the dynamics of biomass and energy use exhibit a temporal trend, but with no detectable difference in the temporal trends for abundance-driven and observed dynamics. However, for a substantial minority of routes (28% overall for biomass; 11% overall for energy use), there is a detectable deviation between the trends expected due only to changes in abundance and the observed dynamics. 

### Table 2.


```{r}

slope_winning_fits <- filter(winning_fits_both, modelcomplexity > 1) %>%
  mutate(abundance_driven_increase = ratio_sim > 1,
         observed_increase = ratio_real > 1)

slope_winning_fits_summary <- slope_winning_fits %>%
  group_by(currency) %>%
  summarize(n_with_slope = dplyr::n(),
            n_abundance_increase = sum(abundance_driven_increase),
            n_observed_increase = sum(observed_increase),
            prop_abundance_increase = round(mean(abundance_driven_increase), digits = 2),
            prop_observed_increase = round(mean(observed_increase), digits = 2))

slope_winning_fits_pretty <- slope_winning_fits_summary %>%
  mutate(currency = ifelse(currency == "biomass", "Total biomass", "Total metabolic flux")) %>%
  select(currency,
         # n_abundance_increase,
         # n_observed_increase,
         prop_abundance_increase,
         prop_observed_increase,
         n_with_slope) %>%
  rename(Currency = currency,
         # `Number of routes with abundance-driven increase`= n_abundance_increase,
         # `Number of routes with observed increase`  = n_observed_increase,
         `Proportion of increasing abundance-driven trends` = prop_abundance_increase,
         `Proportion of increasing observed trends` = prop_observed_increase,
         `Number of routes with temporal trends` = n_with_slope)

slope_winning_fits_pretty

```

#### Table 2. Direction of temporal trends in abundance-driven and observed dynamics.

Restricted to the routes exhibiting temporal trends in total biomass and total metabolic flux, the proportion of trends that are increasing (specifically, the ratio of the last fitted value to the first fitted value > 1) for abundance-driven and observed dynamics. Trends that are not increasing are decreasing.

Trends in abundance-driven dynamics are dominated by *declines* (72% of routes for total biomass, and 76% of routes for total energy). Observed dynamics for biomass differ qualitatively from the abundance-driven dynamics; observed trends in biomass are evenly divided between increases and decreases (50% increasing). Observed trends in energy use more closely mirror abundance-driven trends (69% declines).


\newpage

<!-- # Figure 3. Visualizing decoupling -->

<!-- ```{r} -->

<!-- interaction_all_fits <- all_preds_b %>%  -->
<!--   # get just the change ratios from the interaction models - regardless of whether the interaction model is the best-fitting model. this is because we want an estimate of the slope, we'll color by model fit in a minute -->
<!--   group_by_all() %>%  -->
<!--   filter(grepl( "* source",model_formula), model_family == "Gamma") %>% -->
<!--   mutate(plus = substr(model_formula, 12,12) == "+") %>% filter(!plus) %>% -->
<!--   ungroup() %>% -->
<!--   # now join to *winning* formulas to get the best fitting formula -->
<!--   left_join(select(winning_fits, matssname, model_formula, modelcomplexity) %>% -->
<!--               rename(winning_formula = model_formula)) %>% -->
<!--   select(winning_formula, ratio_sim, ratio_real, matssname, modelcomplexity) %>% -->
<!--   mutate(currency = "Biomass") -->


<!-- interaction_all_fits_energy <- all_preds_e  %>%  -->
<!--   # get just the change ratios from the interaction models - regardless of whether the interaction model is the best-fitting model. this is because we want an estimate of the slope, we'll color by model fit in a minute -->
<!--   group_by_all() %>%  -->
<!--   filter(grepl( "* source",model_formula), model_family == "Gamma") %>% -->
<!--   mutate(plus = substr(model_formula, 12,12) == "+") %>% filter(!plus) %>% -->
<!--   ungroup() %>% -->
<!--   # now join to *winning* formulas to get the best fitting formula -->
<!--   left_join(select(winning_fits_energy, matssname, model_formula, modelcomplexity) %>% -->
<!--               rename(winning_formula = model_formula)) %>% -->
<!--   select(winning_formula, ratio_sim, ratio_real, matssname, modelcomplexity) %>% -->
<!--   mutate(currency = "Metabolic flux") -->


<!-- all_interaction <- bind_rows(interaction_all_fits, interaction_all_fits_energy) %>% -->
<!--   mutate(`Best fitting model` = ifelse(modelcomplexity == 1, "No trend", ifelse(modelcomplexity == 2, "Coupled trend", "Decoupled trends"))) -->


<!-- ``` -->


<!-- ```{r, fig.dim= c(5.5,3)} -->

<!-- decoupling_plot <- ggplot(all_interaction, aes(ratio_sim, ratio_real, color = `Best fitting model`)) +  -->
<!--   geom_point(alpha = .5) +  -->
<!--   #geom_point(data = filter(all_interaction, winning_formula != "1"), aes(color = winning_formula), alpha = .3) +  -->
<!--   scale_y_log10() +  -->
<!--   scale_x_log10() +  -->
<!--   scale_color_viridis_d(option = "mako", end = .8, direction = -1) +  -->
<!--   theme(legend.position = "bottom") +  -->
<!--   geom_abline(intercept = 0, slope = 1) +  -->
<!--  geom_vline(xintercept = 1, linetype = 2) +  -->
<!--   geom_hline(yintercept = 1, linetype =2) + -->
<!--   xlab("Abundance-driven change \n Ratio of last fitted value to first fitted value") +  -->
<!--   ylab("Observed change \n Ratio of last fitted value to first fitted value") + -->
<!--   facet_wrap(vars(currency), ncol = 2) -->


<!-- decoupling_plot -->
<!-- ``` -->


<!-- #### Figure 3. Decoupling between abundance-driven and observed trends. -->

<!-- Observed change (ratio of last fitted value to first fitted value, y-axis) in total biomass (left) and total metabolic flux (right) compared to the change expected only due to changes in total abundance (ratio of last fitted value to first fitted value, x-axis). Values greater than 1 (dashed horizontal and vertical lines) mark positive (increasing) trends, while values less than 1 are negative trends. Each point marks the fitted values from a Gamma log-link generalized linear model of the form `response ~ year * dynamics` for a given route. This estimates separate long-term slopes for observed and abundance-driven dynamics. Points are colored corresponding to the best-fitting model (intercept-only, or "no trend"; a slope for year but no difference in slopes between observed and abundance-driven dynamics, or "coupled trend", and separate slopes for observed and abundance-driven dynamics, "decoupled trends") for each route. Deviations from the 1:1 line (solid black line) reflect changes in the community size structure that modulate the relationship between total abundance and total biomass or energy use.  -->

<!-- Changes in total biomass and total energy use generally track changes driven by fluctuations in total abundance, with appreciable scatter around the 1:1 line. When this translates into a statistically detectable decoupling between observed and abundance-driven dynamics ("Decoupled trends"), this is usually in the form of abundance-driven change being more negative (a steeper decline or a smaller increase) than observed change in biomass or energy use (a less steep decline or larger increase), resulting in points falling above and to the left of the 1:1 line. This occurs more strongly and frequently for biomass than for metabolic flux. -->


<!--  \newpage -->
 
# Appendix S1 Figure S2

```{r}


winning_fits <- winning_fits %>% 
  left_join(all_cs_compares) %>%
  mutate(Syndrome = ifelse(modelcomplexity == 1, "No trend", ifelse(modelcomplexity == 2, "Coupled trend", "Decoupled trends")))


winning_fits <- winning_fits %>%
  mutate(categorical_fit = Syndrome) %>%
  mutate(categorical_fit = as.factor(categorical_fit)) %>%
  mutate(abs_log_ratio = abs(log(real_end_mean_mass / real_begin_mean_mass)),
         log_ratio = log(real_end_mean_mass/real_begin_mean_mass)) %>%
  mutate(isd_turnover = 1 - real_overlap)

isd_mean_change_plot <-  ggplot(winning_fits, aes(abs_log_ratio, color = Syndrome, fill= Syndrome)) + 
  geom_histogram() + 
  facet_wrap(vars(Syndrome), scales = 'free_y', ncol = 1) +
  scale_color_viridis_d(option = "mako", end = .8, direction = -1) +     
  scale_fill_viridis_d(option = "mako", end = .8, direction = -1) +
  theme(legend.position = "none") +
  xlab("Mean mass absolute log ratio") +
  ylab("Count") +
  ggtitle("Mean size change")
# 
# mean_change_coupled <- filter(winning_fits, modelcomplexity == 2)
# mean_change_decoupled <- filter(winning_fits, modelcomplexity == 4)
# mean_change_nt <- filter(winning_fits, modelcomplexity == 1)
# 
# ks.test(mean_change_coupled$ratio_real, mean_change_decoupled$ratio_real)
# 
# 
# ks.test(mean_change_coupled$real_overlap, mean_change_decoupled$real_overlap)
# 
# ks.test(mean_change_coupled$real_overlap, mean_change_nt$real_overlap)
# 
# 
# ks.test(mean_change_decoupled$real_overlap, mean_change_nt$real_overlap)
# 
# ks.test(c(mean_change_nt$real_overlap,mean_change_coupled$real_overlap), mean_change_decoupled$real_overlap)


```

```{r}

isd_overlap_plot <- ggplot(winning_fits, aes(isd_turnover,color = Syndrome, fill= Syndrome)) + 
  geom_histogram() + 
  facet_wrap(vars(Syndrome), scales = 'free_y', ncol = 1) +
  scale_color_viridis_d(option = "mako", end = .8, direction = -1) +  scale_fill_viridis_d(option = "mako", end = .8, direction = -1) +
  theme(legend.position = "none") +
  xlab("ISD turnover") +
  ylab("Count") +
  ggtitle("ISD turnover")




bcd_plot <- ggplot(winning_fits, aes(sp_turnover_bcd,color = Syndrome, fill= Syndrome)) + 
  geom_histogram() + 
  facet_wrap(vars(Syndrome), scales = 'free_y', ncol = 1) +
  scale_color_viridis_d(option = "mako", end = .8, direction = -1) +  scale_fill_viridis_d(option = "mako", end = .8, direction = -1) +
  theme(legend.position = "none") +
  xlab("Bray-Curtis index") +
  ylab("Count") +
  ggtitle("Species turnover")
 # isd_overlap_plot
# 
# isd_glm <- glm(real_overlap ~ model_formula, family = "binomial", data = winning_fits)
# #
# # AIC(isd_glm)
# # AIC(glm(real_overlap ~ 1, family = "binomial", data = winning_fits))
# 
# glm_cor <- cor(predict(isd_glm, type = "response"), isd_glm$model[,1]) ^ 2
```

```{r, fig.dim = c(6, 3.5)}

isd_change_plots <- multi_panel_figure(columns = 3, rows = 1) %>%
  fill_panel(isd_mean_change_plot) %>%
  fill_panel(isd_overlap_plot) %>%
  fill_panel(bcd_plot)
isd_change_plots
```


#### Figure 4. Histograms of change in community structure for routes showing no trends, coupled, and decoupled trends in abundance-driven and observed dynamics.

Histograms of (A) change in mean body size from the first to the last five years of monitoring, (B) overall change in the size structure, and (C) change in species composition for routes whose dynamics for  total biomass were best-described using no temporal trend (bottom row; intercept-only model), separate trends for observed and abundance-driven dynamics (middle row), or the same trend for observed and abundance-driven dynamics (top row). Change in mean body size (A) is calculated as the ratio of the mean body size of all individuals observed in the last 5 years of the timeseries relative to the mean body size of all individuals observed in the first 5 years. Overall change in the ISD (B) is calculated as the degree of turnover between the ISDs for the first and last five years of the timeseries (see text). Change in species composition (C) is Bray-Curtis dissimilarity comparing species composition in the first five years to the last five years. 

Routes that exhibit decoupling between observed and abundance-driven changes in total biomass exhibit a high prevalence of increases and decreases in mean body size (middle row, panel A) compared to the changes seen in routes that show either no trend or non-decoupled trends. However, routes with all three signatures of dynamics (coupling, decoupling, or no trend) are not detectably different in the degree of overall change in the ISD or in species composition over time (panels B and C).

\newpage

# Statistical comparisons of distributions in Figure S2

### Mean mass

#### Appendix S1 Table S3. 

```{r, results = T}

mean_mass_lm <- lm(abs_log_ratio ~ categorical_fit, data = winning_fits)

mean_mass_lm_intercept <- lm(abs_log_ratio ~ 1, data = winning_fits)

anova(mean_mass_lm, mean_mass_lm_intercept)

```


ANOVA comparing the models `abs_log_ratio_mean_mass ~ best fitting model type` and `abs_log_ratio_mean_mass ~ 1`. The fit incorporating model type is superior to the intercept-only model (p < 0.0001).

#### Table 4. Model estimates for absolute log ratio of mean mass for routes best-described by different dynamics.

```{r, results = T}
mean_mass_lm_emmeans <- emmeans::emmeans(mean_mass_lm, specs = ~ categorical_fit)

as.data.frame(mean_mass_lm_emmeans)
```

Estimates (calculated using `emmeans` [@lenth2021]) for the mean absolute log ratio of mean mass for routes whose dynamics for biomass best-described by different model types. Routes with decoupled long-term trends between biomass and abundance-driven dynamics have higher absolute log ratios (mean .56, 95% credible interval .53-.58) than routes with covarying trends in biomass and abundance (mean of .2; 95% interval .18-.22) or no detectable temporal trend (mean of .22; .2-.24). 

#### Table 5. Contrasts for absolute log ratio of mean mass.

```{r, results = T}
as.data.frame(pairs(mean_mass_lm_emmeans))
```

Contrasts for the above comparisons. There is a significant contrast between routes with decoupled trends and the other two types of dynamics (both contrasts, p < 0.001), but not between "no trend" and "coupled trend" routes (contrast p = .31).


### ISD turnover

#### Table 6. ANOVA for turnover in the ISD


```{r, results = T}

overlap_glm <- glm(isd_turnover ~ categorical_fit, data = winning_fits, family = "binomial")

overlap_glm_intercept <- glm(isd_turnover ~ 1, data = winning_fits, family = "binomial")

anova(overlap_glm, overlap_glm_intercept, test = "Chi")

```

### Species compositional turnover

#### Table 7. ANOVA for Bray-Curtis dissimilarity


```{r, results = T}

bcd_glm <- glm(sp_turnover_bcd ~ categorical_fit, data = winning_fits, family = "binomial")

bcd_glm_intercept <- glm(sp_turnover_bcd ~ 1, data = winning_fits, family = "binomial")

anova(bcd_glm, bcd_glm_intercept, test = "Chi")

```


ANOVA comparing a binomial GLM of the form `bray curtis dissimilarity ~ best fitting model type` to an intercept-only model. The `best fitting model type` model is not superior to the intercept only model (p = .37).



\newpage
# References
